--- 
title: "Assignment 1.2 - R Refresher"
author: "Ragunath Gunasekaran"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Import, Plot, Summarize, and Save Data  and summary for every variable, structure and type of data elements


```{r RDataLoadInstallPakcages, echo=FALSE, include=FALSE}
# Load Libraries
# blscrapeR: An API Wrapper for the Bureau of Labor Statistics (BLS)
library(blscrapeR)  
library(pastecs)
library(ggplot2)

## Load Data, Read File 
##setwd("C:\Users\ragun\Documents\GitHub\DSC630")

bls_2000 <- bls_api(c("LNS13327709", "LNS14000000"),
                    startyear = 1993, endyear = 2000)

bls_2008 <- bls_api(c("LNS13327709", "LNS14000000"),
                    startyear = 2001, endyear = 2008)

bls_2016 <- bls_api(c("LNS13327709", "LNS14000000"),
                    startyear = 2009, endyear = 2016)



bls_2020 <- bls_api(c("LNS13327709", "LNS14000000"),
                    startyear = 2016, endyear = 2020)

# Dataframe Checks
#head(bls_2000,20)
#str(bls_2000)
#summary(bls_2000)
#class(bls_2000)
#dim(bls_2000)

```


The below is the Structure of datasets which explains the detail of metadata.

```{r RMetadata, echo=FALSE}
summary(bls_2000$value)
summary(bls_2008$year)
summary(bls_2016$periodName)


str(bls_2000)
str(bls_2008)
str(bls_2016)

dim(bls_2000)
dim(bls_2008)
dim(bls_2016)

#write.csv(bls_2000, "2000Data.csv")
#write.csv(bls_2008, "2008Data.csv")
#write.csv(bls_2016, "2016Data.csv")

```


The Dataset contains 180 rows and 6 Variables in 2000 data


## 1. Histograms, Box plots

```{r plots, echo=FALSE}

ggplot(data = bls_2008, mapping = aes(x = periodName, y = value,colour=periodName)) +
  geom_boxplot()+ggtitle("Box Plot for Data 2008")+labs(y="value", x = "periodName")

ggplot(bls_2016, aes(x=value)) + 
  geom_histogram(color="red", fill="blue")+ggtitle("Histogram Plot for 2016 SR")+labs(y="Freq of Values", x = "values")

ggplot(data = bls_2000, mapping = aes(x = value, colour = year)) +
  geom_freqpoly(binwidth = 0.1)+ggtitle("Freqpoly Plot for 2000")+labs(y="Freq of values", x = "value")

```


```{r summaryanalytis, echo=FALSE}
summary(bls_2000)
summary(bls_2008)
summary(bls_2016)
summary(bls_2020)

```


### 2. Explore Some Bivariate Relations

```{r correlation, echo=FALSE}
library(tidyverse)

cor(bls_2008$value, bls_2008$year)
cor(bls_2016$value, bls_2016$year)
cor(bls_2020$value, bls_2020$year)

```


### 3. Organize a Data Report
```{r writedata}

summary(bls_2000)
summary(bls_2008)

```



# Conclusion - Summary of Analysis

As part of the analysis, we have compared US Unemployment Rate from 2000 to 2020 with various period.



# References:

1. https://cran.r-project.org/web/packages/blscrapeR/index.html

2. http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization


